---
title: "a3_task3_caroline_shepherd"
author: "Caroline Shepherd"
date: "10/27/2021"
output: 
  html_document: 
    toc: yes
    theme: cerulean
    code_folding: hide
---
## Overview

This code wrangles and visualizes data regarding the size of sea urchins, as it relates to species.

Data Source: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

## Setup

```{r setup, include=FALSE, warnings = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Attach packages

```{r}
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(dplyr)
library(lubridate)
```

## Read in the data

```{r}
urchins <- read_csv(here("data", "urchins.csv"))

## Clean names

urchins_clean <- urchins %>% 
  clean_names()
```

## Visualize the data

```{r}
ggplot(data = urchins_clean, aes(x = size)) + # Create ggplot
  geom_histogram(bins = 10, aes(fill = common_name)) + # Change number of bins from default to 10, and fill color to common name column
  facet_wrap(~ common_name) + # Facet wrap by common name to sort Purple and Red Sea Urchins
  scale_fill_manual(values = c("#800080", "#FF0000")) + # Change colors of bins
  labs(x = "Number of Urchins", # Change axes titles
       y = "Size of Urchins (cm)",
       title = "Size (cm) and Amount of Purple and Red Sea Urchins",
       caption = "Figure 1.0 depicts the size (in cm) and the amount of sea urchins, sorted by species (Red or Purple Sea Urchins)") +
  theme(plot.caption = element_text(hjust = 0.5), # Center the figure caption
        plot.title = element_text(hjust = .5)) # Center the figure title
```

#### Table 1.0 Depicts the statistical analysis of Sea Urchin Size (cm), grouped by species of Red or Purple Sea Urchins
```{r}
urchin_stats <- urchins_clean %>%  # Name new vector
  group_by(common_name) %>%  # Group by species
  summarize(mean_urchins = round(mean(size, na.rm = TRUE),2), # Use summarize to find several statistics 
            median_urchins = round(median(size, na.rm = TRUE),2),
            sd_urchins = round(sd(size, na.rm = TRUE), 2),
            sample_size = n()) %>% 
   rename("Common Name" = common_name, # Rename table names for clarity
         "Mean Sea Urchin Size (cm)" = mean_urchins,
         "Median Sea Urchin Size (cm)" = median_urchins,
         "Standard Deviation of Sea Urchin Size (cm)" = sd_urchins,
         "Amount of Sea Urchins Sampled" = sample_size)

urchin_stats %>% # Add caption to chart
  kable(caption = "Table 1.0 Statistical Analysis of Sea Urchin Size (cm), Grouped by Species")
  
```

### The table and graph above explore the size differences among Red Sea Urchins and Purple Sea Urchins. The mean and median size of Red Sea Urchins is larger than the mean and median size of Purple Sea Urchins for the populations included in this data. The spread of Purple Sea Urchin size is slightly positively skewed, and the distribution of Red Sea Urchins is normal. Red Sea Urchins have a 33% larger mean size as compared to the Purple Sea Urchins.

## Purple Sea Urchin Data Exploration
```{r}
purple_urchins <- urchins_clean %>% 
  filter(common_name == "Purple Urchin") %>% # Only include Purple Sea Urchins
  mutate(date_new = mdy(date)) %>%  # Add new date class
  mutate(year = year(date_new)) %>%  # Add new column for year
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpenteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"))
```


## Purple Sea Urchin Summary Table
```{r}
purple_urchins_sum <- purple_urchins %>% 
  group_by(site_full, year) %>% # Group by site and by year
  summarize(count = n()) # Summarize number of purple urchins based on site and year
```

## Make graph to show the amount of Purple Sea Urchins at each location

```{r}
ggplot(data = purple_urchins_sum, aes(x = year, y = count)) + # Plot the data with x and y axes
  geom_line(aes(color = site_full)) + # Fill color by site
  theme_bw() + # Change the theme to black and white
  labs(x = "Year", # Update title and axes titles
       y = "Amount of Purple Sea Urchins",
       title = "Amount of Purple Sea Urchins (per year) Per Site",
       caption = "Figure 2.0: This graph depicts the amount of Purple Sea Urchins present at each location listed, and is shown over time (in years)",
       color = "Site Location") +
  theme(plot.caption = element_text(hjust = 0.2))
  
```

## Figure 2.0 reflects the changes in purple sea urchin populations for these locations. The graph shoes that the purple sea urchin population is in decline, but there are other factors to take into consideration, like human error. Less sampling events could lead to a smaller count of urchins, which would make the graph look like the populations are declining when they are not declining in reality.


